---
title: "L. goldeii and L. fus telemetry paper 2015-Apr 2017"
author: "Ajay Venkataraman"
date: "2024-07-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Install packages to add and manipulate data

```{r packages}
library(tidyverse)
library(ggmap)
library(VTrack)
library(leaflet)
library(openxlsx)
library(data.table)
library(htmlwidgets)
library(sp)
getwd()
```
# Import csv files to R (change filepath if need be)


```{r }

bb1 <- read_csv("../data/1.BLACK BASS.csv")
bb2 <- read_csv("../data/1.1.wet seaon 2016-17.csv")
bb3 <- read_csv("../data/2.Baia 2017.csv")
bb4 <- read_csv("../data/2.1.Baia sep 2018 last download this study.csv")
```
# Merge and first clean up on telemetry dataset

```{r }
df1 <- bind_rows(bb1, bb2, bb3, bb4) %>%
    select(!c("Transmitter Name", "Transmitter Serial", "Sensor Value", "Sensor Unit", "Transmitter Type", "Sensor Precision"))
```

## Split transmitter and receiver column values to retain only second numbers

```{r }
df1$Receiver <- sapply(strsplit(df1$Receiver, "-"), function(x) x[2])
df1$Transmitter <- sapply(strsplit(df1$Transmitter, "-"), function(x) tail(x, 1))
```

## Change time from UTC to AEST and to POSIXCT format
```{r }
df1 <- 
  df1 %>%
  mutate(datetimeEST = with_tz(as.POSIXct(`Date and Time (UTC)`, format = "%Y-%m-%d %H:%M:%S",  tz= "UTC")), tzone = "Australia/Brisbane", .keep = "unused") %>% ## Change time from UTC to AEST and to POSIXCT format 
  select(!tzone) ##remove timezone column
```

# Import and clean tagged fish and receiver data
```{r }
fishtags <- read_csv("../data/tagged fish 2015-2017.csv")
fishtags <- fishtags[ , -c(8:22)]
```

#Import receiver history data (REDUNDANT)
```{r}
receivers <- read.csv("C:/Users/jc347879/OneDrive - James Cook University/Documents/Ajay PhD/writing/thesis_chapter 2/ajay sheets/receivers history jun2015-apr2017.csv")
receivers <- receivers[-c(41:102), ]
receivers [receivers == ""] <- NA
## Convert date columns to the correct format
receivers$dd1 <- as.POSIXct(receivers$dd1)
receivers$dr1 <- as.POSIXct(receivers$dr1)
receivers$dd2 <- as.POSIXct(receivers$dd2)
receivers$dr2 <- as.POSIXct(receivers$dr2)
receivers$dd3 <- as.POSIXct(receivers$dd3)
receivers$dr3 <- as.POSIXct(receivers$dr3)
receivers$dd4 <- as.POSIXct(receivers$dd4)
receivers$dr4 <- as.POSIXct(receivers$dr4)
```

# Merge tagged fish with telemetry dataset
```{r}
df2 <- merge(df1, fishtags, by.x = "Transmitter", by.y = "transmitter", all.x = TRUE)

df1A <- rename(df1, transmitter=Transmitter)
fishtags$transmitter <- as.character(fishtags$transmitter)
df2A <- df1A %>% 
  left_join(fishtags, by= "transmitter")
```

# Create filtered df2: Filter df2 to include observations only till 1st of April 2017
```{r}
target_date <- as.POSIXct("2017-04-01", tz = "EST")
filtered_df2 <- df2 %>%
  filter(complete.cases(species), species != "", datetimeEST <= target_date)
```

# Clean up filtered_df2: Rename date column to tagged date for species
```{r}
filtered_df2 <- rename(filtered_df2, tagdate = date)
## Split datetimeEST to date
filtered_df2 <- filtered_df2 %>% mutate(date=date(datetimeEST)) 
## Keep only non-blank species rows
filtered_df2 <- filtered_df2 %>%
  filter(!is.na(species) & species != "")
```

# Ensure datetimeEST is in the correct date-time format
```{r}
filtered_df2$datetimeEST <- as.POSIXct(filtered_df2$datetimeEST, format="%Y-%m-%d %H:%M:%S")
```

# Use case_when to create the Station.Name column based on updated conditions
```{r}
filtered_df2 <- filtered_df2 %>%
  mutate(Station.Name = case_when(
    # LL1
    Receiver == "125667" & datetimeEST >= as.POSIXct("2015-06-18") & datetimeEST <= as.POSIXct("2016-11-22") ~ "LL1",
    Receiver == "127283" & datetimeEST >= as.POSIXct("2016-11-21") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL1",
    
    # LL11
    Receiver == "125668" & datetimeEST >= as.POSIXct("2015-06-18") & datetimeEST <= as.POSIXct("2015-11-09") ~ "LL11",
    
    # LL14
    Receiver == "125669" & datetimeEST >= as.POSIXct("2015-06-20") & datetimeEST <= as.POSIXct("2015-11-10") ~ "LL14",
    Receiver == "125669" & datetimeEST >= as.POSIXct("2016-03-31") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL14",
    
    # LL15
    Receiver == "125670" & datetimeEST >= as.POSIXct("2015-06-18") & datetimeEST <= as.POSIXct("2015-11-10") ~ "LL15",
    Receiver == "127277" & datetimeEST >= as.POSIXct("2016-03-31") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL15",
    
    # LL2
    Receiver == "125671" & datetimeEST >= as.POSIXct("2015-06-18") & datetimeEST <= as.POSIXct("2016-03-30") ~ "LL2",
    Receiver == "127286" & datetimeEST >= as.POSIXct("2016-03-31") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL2",
    
    # LL3
    Receiver == "125672" & datetimeEST >= as.POSIXct("2015-06-18") & datetimeEST <= as.POSIXct("2016-11-22") ~ "LL3",
    Receiver == "127289" & datetimeEST >= as.POSIXct("2016-11-21") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL3",
    
    # LL4
    Receiver == "125673" & datetimeEST >= as.POSIXct("2015-06-18") & datetimeEST <= as.POSIXct("2016-11-20") ~ "LL4",
    Receiver == "127271" & datetimeEST >= as.POSIXct("2016-11-21") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL4",
    
    # LL7
    Receiver == "125674" & datetimeEST >= as.POSIXct("2015-06-19") & datetimeEST <= as.POSIXct("2016-11-20") ~ "LL7",
    Receiver == "127272" & datetimeEST >= as.POSIXct("2016-11-21") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL7",
    
    # LL9
    Receiver == "125675" & datetimeEST >= as.POSIXct("2015-06-19") & datetimeEST <= as.POSIXct("2016-11-21") ~ "LL9",
    Receiver == "127279" & datetimeEST >= as.POSIXct("2016-11-22") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL9",
    
    # LL8
    Receiver == "127273" & datetimeEST >= as.POSIXct("2015-06-23") & datetimeEST <= as.POSIXct("2016-11-21") ~ "LL8",
    Receiver == "127290" & datetimeEST >= as.POSIXct("2016-11-22") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL8",
    
    # LL6
    Receiver == "127271" & datetimeEST >= as.POSIXct("2015-06-23") & datetimeEST <= as.POSIXct("2016-11-19") ~ "LL6",
    Receiver == "125670" & datetimeEST >= as.POSIXct("2016-11-20") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL6",
    
    # LL5
    Receiver == "127272" & datetimeEST >= as.POSIXct("2015-06-23") & datetimeEST <= as.POSIXct("2016-11-19") ~ "LL5",
    Receiver == "127284" & datetimeEST >= as.POSIXct("2016-11-20") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL5",
    
    # LL13
    Receiver == "127283" & datetimeEST >= as.POSIXct("2016-03-31") & datetimeEST <= as.POSIXct("2016-11-17") ~ "LL13",
    
    # LL12
    Receiver == "127282" & datetimeEST >= as.POSIXct("2015-07-12") & datetimeEST <= as.POSIXct("2015-11-09") ~ "LL12",
    Receiver == "127281" & datetimeEST >= as.POSIXct("2016-03-31") & datetimeEST <= as.POSIXct("2016-11-17") ~ "LL12",
    
    # LL10
    Receiver == "127288" & datetimeEST >= as.POSIXct("2015-10-03") & datetimeEST <= as.POSIXct("2016-11-20") ~ "LL10",
    Receiver == "127280" & datetimeEST >= as.POSIXct("2016-11-21") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL10",
    
    # P1
    Receiver == "125678" & datetimeEST >= as.POSIXct("2015-06-20") & datetimeEST <= as.POSIXct("2015-11-08") ~ "P1",
    Receiver == "125678" & datetimeEST >= as.POSIXct("2016-04-08") & datetimeEST <= as.POSIXct("2016-09-26") ~ "P1",
    Receiver == "127292" & datetimeEST >= as.POSIXct("2016-09-27") & datetimeEST <= as.POSIXct("2016-11-17") ~ "P1",
    
    # P2
    Receiver == "125676" & datetimeEST >= as.POSIXct("2015-06-19") & datetimeEST <= as.POSIXct("2015-11-08") ~ "P2",
    Receiver == "125676" & datetimeEST >= as.POSIXct("2016-04-08") & datetimeEST <= as.POSIXct("2016-09-26") ~ "P2",
    Receiver == "127266" & datetimeEST >= as.POSIXct("2016-09-27") & datetimeEST <= as.POSIXct("2016-11-17") ~ "P2",
    
    # P3
    Receiver == "125677" & datetimeEST >= as.POSIXct("2015-06-19") & datetimeEST <= as.POSIXct("2015-11-08") ~ "P3",
    Receiver == "125677" & datetimeEST >= as.POSIXct("2016-04-08") & datetimeEST <= as.POSIXct("2016-11-18") ~ "P3",
    
    # P4
    Receiver == "127264" & datetimeEST >= as.POSIXct("2015-06-20") & datetimeEST <= as.POSIXct("2015-11-07") ~ "P4",
    Receiver == "127264" & datetimeEST >= as.POSIXct("2016-04-08") & datetimeEST <= as.POSIXct("2016-11-17") ~ "P4",
    
    # LL-P
    Receiver == "127282" & datetimeEST >= as.POSIXct("2015-11-09") & datetimeEST <= as.POSIXct("2016-11-21") ~ "LL-P",
    Receiver == "127281" & datetimeEST >= as.POSIXct("2016-11-22") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LL-P",
    
    # T8
    Receiver == "127266" & datetimeEST >= as.POSIXct("2015-06-22") & datetimeEST <= as.POSIXct("2015-11-09") ~ "T8",
    Receiver == "127290" & datetimeEST >= as.POSIXct("2016-04-07") & datetimeEST <= as.POSIXct("2016-11-18") ~ "T8",
    
    # T6
    Receiver == "127268" & datetimeEST >= as.POSIXct("2015-06-22") & datetimeEST <= as.POSIXct("2015-11-10") ~ "T6",
    
    # T3
    Receiver == "127267" & datetimeEST >= as.POSIXct("2015-06-22") & datetimeEST <= as.POSIXct("2015-11-10") ~ "T3",
    
    # T2
    Receiver == "127270" & datetimeEST >= as.POSIXct("2015-06-22") & datetimeEST <= as.POSIXct("2015-11-10") ~ "T2",
    
    # T4
    Receiver == "127269" & datetimeEST >= as.POSIXct("2015-06-22") & datetimeEST <= as.POSIXct("2015-11-10") ~ "T4",
    Receiver == "127267" & datetimeEST >= as.POSIXct("2016-04-07") & datetimeEST <= as.POSIXct("2016-11-19") ~ "T4",
    
    # T1
    Receiver == "127263" & datetimeEST >= as.POSIXct("2015-06-30") & datetimeEST <= as.POSIXct("2015-11-09") ~ "T1",
    
    # T5
    Receiver == "127275" & datetimeEST >= as.POSIXct("2015-07-10") & datetimeEST <= as.POSIXct("2015-11-10") ~ "T5",
    
    # LLC2
    Receiver == "127277" & datetimeEST >= as.POSIXct("2015-07-12") & datetimeEST <= as.POSIXct("2015-11-10") ~ "LLC2",
    Receiver == "125670" & datetimeEST >= as.POSIXct("2015-11-10") & datetimeEST <= as.POSIXct("2016-11-18") ~ "LLC2",
    Receiver == "545562" & datetimeEST >= as.POSIXct("2016-11-19") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LLC2",
    
    # LLC1
    Receiver == "127278" & datetimeEST >= as.POSIXct("2015-07-12") & datetimeEST <= as.POSIXct("2015-11-09") ~ "LLC1",
    Receiver == "545563" & datetimeEST >= as.POSIXct("2016-11-19") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LLC1",
    
    # LLC3
    Receiver == "127279" & datetimeEST >= as.POSIXct("2015-07-12") & datetimeEST <= as.POSIXct("2016-11-16") ~ "LLC3",
    Receiver == "127264" & datetimeEST >= as.POSIXct("2016-11-17") & datetimeEST <= as.POSIXct("2017-04-01") ~ "LLC3",
    
    # OPC
    Receiver == "127281" & datetimeEST >= as.POSIXct("2015-07-12") & datetimeEST <= as.POSIXct("2016-11-09") ~ "OPC",
    Receiver == "127280" & datetimeEST >= as.POSIXct("2015-11-10") & datetimeEST <= as.POSIXct("2016-11-18") ~ "OPC",
    Receiver == "127266" & datetimeEST >= as.POSIXct("2016-11-19") & datetimeEST <= as.POSIXct("2017-04-01") ~ "OPC",
    
    # PC1
    Receiver == "127280" & datetimeEST >= as.POSIXct("2015-07-12") & datetimeEST <= as.POSIXct("2015-11-06") ~ "PC1",
    Receiver == "545562" & datetimeEST >= as.POSIXct("2015-11-07") & datetimeEST <= as.POSIXct("2016-11-18") ~ "PC1",
    Receiver == "545561" & datetimeEST >= as.POSIXct("2016-09-25") & datetimeEST <= as.POSIXct("2017-04-01") ~ "PC1",
    
    # TR1
    Receiver == "127289" & datetimeEST >= as.POSIXct("2015-09-30") & datetimeEST <= as.POSIXct("2016-09-20") ~ "TR1",
    
    # TR2
    Receiver == "127285" & datetimeEST >= as.POSIXct("2015-09-30") & datetimeEST <= as.POSIXct("2016-09-20") ~ "TR2",
    
    # RDP
    Receiver == "127287" & datetimeEST >= as.POSIXct("2016-04-08") & datetimeEST <= as.POSIXct("2016-09-24") ~ "RDP",
    Receiver == "127292" & datetimeEST >= as.POSIXct("2016-09-19") & datetimeEST <= as.POSIXct("2017-04-01") ~ "RDP",
    
    # RLL
    Receiver == "127292" & datetimeEST >= as.POSIXct("2016-04-08") & datetimeEST <= as.POSIXct("2016-09-26") ~ "RLL",
    Receiver == "127287" & datetimeEST >= as.POSIXct("2016-09-25") & datetimeEST <= as.POSIXct("2017-04-01") ~ "RLL",
    
    # T9
    Receiver == "127276" & datetimeEST >= as.POSIXct("2015-07-12") & datetimeEST <= as.POSIXct("2015-09-30") ~ "T9",
    
    # Default
    TRUE ~ NA_character_
  ))
```

# Add coordinates for receiver sites
```{r}
station_coords <- data.frame(
  Station.Name = c("LL1", "LL11", "LL14", "LL15", "LL2", "LL3", "LL4", "LL7", "LL9", "LL8",
                   "LL6", "LL5", "LL13", "LL12", "LL10", "P1", "P2", "P3", "P4", "LL-P",
                   "LLC2", "LLC1", "LLC3", "OPC", "PC1", "T1", "T2", "T3", "T4", "T5", "T6", "T7", "T8"),
  Latitude = c(-4.95223, -4.95115, -4.96095, -4.96123, -4.95548, -4.95937, -4.95283, -4.94506, -4.94601, -4.94356,
               -4.94692, -4.95043, -4.95967, -4.95383, -4.94987, -4.93681, -4.94228, -4.94654, -4.95575, -4.94918,
               -4.94625, -4.94637, -4.94773, -4.93797, -4.93293, -4.71193, -4.71568, -4.71802, -4.72402, -4.71747, -4.71335, -4.71627, -4.71522),
  Longitude = c(151.4746, 151.4851, 151.4897167, 151.4889667, 151.47539, 151.47065, 151.4689, 151.45265, 151.44564, 151.45083,
                151.4588, 151.46561, 151.4886, 151.48956, 151.44245, 151.4434, 151.4392667, 151.43913, 151.42575, 151.4421333,
                151.4784667, 151.4821167, 151.4749667, 151.45551, 151.4510333, 151.65758, 151.66127, 151.66362, 151.6689, 151.67103, 151.68275, 151.68728, 151.70337)
)
```

## Join coordinates data frame to filtered_df2 data frame
```{r}
filtered_df2 <- filtered_df2 %>%
  left_join(station_coords, by = "Station.Name")
```

## Remove extra columns and rename latlong columns
```{r}
filtered_df2 <- filtered_df2[ -c(5,6)]
filtered_df2 <- rename(filtered_df2, lat = Latitude.y)
filtered_df2 <- rename(filtered_df2, long = Longitude.y)
```

```{r}
# Isolate by species IF NEED BE and create CSV
lg5 <- filter(filtered_df2, species=="lg5")
write.csv(lg13, file = "C:/Users/jc347879/OneDrive - James Cook University/Documents/Ajay PhD/writing/thesis_chapter 2/ajay sheets/lg13.csv")
```

# PLOTTING

## Create an object for the species to plot for
```{r}
species_to_plot <- "lg22"
## Abacus plot 
filtered_df2 %>%
     filter(species == species_to_plot) %>%
     ggplot(mapping = aes(x = datetimeEST, y = as.factor(Station.Name))) +
     xlab("Date") + ylab("Receiver station") +
     geom_point() +
     ggtitle(paste("Plot for species:", species_to_plot))
```

## Spatial mapping (register google key)
```{r}
register_google(key='AIzaSyA3i5Enp3gOTjgh8XzqxlcO6zeIXiOxHJo')
```

# superimpose receiver sites on map
```{r}
baia_map <- leaflet() %>%
  # Base group
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
addCircles(lng = station_coords$Longitude, 
           lat = station_coords$Latitude, 
           fill=TRUE, color="grey", weight=10, opacity=0.8)
baia_map
```

# arrange the dataframe by different species
```{r}
filtered_df2 <- filtered_df2 %>% 
  arrange(species, Station.Name, datetimeEST)
```

# create new dataframes for each species
```{r}
species_list <- split(filtered_df2, filtered_df2$species)
```

# create new excel files for each species
```{r}
for (species_name in names(species_list)) {
     file_name <- paste0("/Users/jc347879/OneDrive - James Cook University/Documents/Ajay PhD/writing/thesis_chapter 2/species detections", species_name, ".xlsx")
     write.xlsx(species_list[[species_name]], file = file_name)
 }
```

# station count per species
```{r}
station_count_per_species <- filtered_df2 %>%
  group_by(species) %>%
  summarise(Unique_Stations = n_distinct(Station.Name))
```

# Calculate the number of detections per species
```{r}
detections_per_species <- filtered_df2 %>%
  group_by(species) %>%
  summarise(Detections_Per_Species = n())
```

# Merge the detections information with the station count per species dataframe
```{r}
station_count_per_species <- station_count_per_species %>%
  left_join(detections_per_species, by = "species")
```  
# MEAN RESIDENCY INDEX BY SPECIES

# sort the dataframe by species, station name and datetime
```{r}
filtered_df2 <- filtered_df2 %>% arrange(species, Station.Name, datetimeEST)
```

# Calculate the time difference between consecutive detections for the same species at the same station
```{r}
filtered_df2 <- filtered_df2 %>%
  group_by(species, Station.Name) %>%
  mutate(Time_Diff = as.numeric(difftime(lead(datetimeEST), datetimeEST, units = "hours")))
```

# Replace NA values in Time_Diff with 0
```{r}
filtered_df2$Time_Diff[is.na(filtered_df2$Time_Diff)] <- 0
```

# Calculate the residency index based on time spent
```{r}
filtered_df2 <- filtered_df2 %>%
  mutate(Residency_Index = 1 / Time_Diff)
```

# Change infinite values to zeroes
```{r}
filtered_df2$Residency_Index[is.infinite(filtered_df2$Residency_Index)] <- 0  
```

# Calculate the mean residency index per species
```{r}
mean_residency_per_species <- filtered_df2 %>%
  group_by(species) %>%
  summarise(Mean_Residency_Index = mean(Residency_Index, na.rm = TRUE))
```

# Calculate the mean residency index per station
```{r}
mean_residency_per_station <- filtered_df2 %>%
  group_by(Station.Name) %>%
  summarise(Mean_Residency_Index = mean(Residency_Index, na.rm = TRUE))
```
# MEAN RESIDENCE

# Calculate the mean residence time per station for each species
```{r}
mean_residence_per_station <- filtered_df2 %>%
  group_by(Station.Name) %>%
  summarise(Mean_Residence_Time = mean(Time_Diff, na.rm = TRUE))
```
